---
title: "Exploring OJS"
format: 
  closeread-html:
    code-tools: true
    cr-style:
      narrative-background-color-overlay: "#111111dd"
      narrative-text-color-overlay: white
      narrative-background-color-sidebar: transparent
      section-background-color: transparent
---

Since I'm more comfortable with R, I'll import and clean the data with R.

I can then pass the data frame to `ojs` with `ojs_define()`

```{r}
#| output: false
library(tidyverse, quietly = TRUE)
library(gt)

#list of all constructors
Constructors <- read_csv('data/constructors.csv', show_col_types = FALSE) 

#results by constructor
ConstructorsResults <- read_csv(file = 'data/constructor_results.csv', show_col_types = FALSE) %>%
  left_join(select(Constructors, constructorId, name), by = 'constructorId')

#list of all races  
Races <- read_csv('data/races.csv', show_col_types = FALSE)

#add race info to constructors results
ConstructorsResults <- left_join(ConstructorsResults, 
                                 Races %>% select(raceId, year, round, name, date), 
                                 by = 'raceId') 
#Create list of 2024 results
Constructors2024 <- ConstructorsResults %>%
  filter(year > 2023) %>%
  select(-c(constructorResultsId,
            raceId, 
            constructorId,
            status)) %>%
  rename(team = name.x,
         race = name.y) %>%
  select(round, date, race, team, points, year) %>%
  arrange(round, desc(points)) %>%
  group_by(team) 

Mclaren_col     <- '#ff8000'
Ferrari_col     <- '#e80020'
RedBull_col     <- '#3671c6'
Mercedes_col    <- '#27f4d2'
AstonMartin_col <- '#229971'
Alpine_col      <- '#ff36a1'
Haas_col        <- '#ffffff'
RB_col          <- '#6692ff'
Williams_col    <- '#64c4ff'
Kick_col        <- '#00e700'

Constructors2024 <- Constructors2024 %>%
  #create a standings column and assign team colours for use in visuals
  mutate(standings = cumsum(points),
         colour = case_when(team == 'Red Bull' ~ RedBull_col,
                            team == 'Ferrari' ~ Ferrari_col,
                            team == 'Mercedes' ~ Mercedes_col,
                            team == 'McLaren' ~ Mclaren_col,
                            team == 'Aston Martin' ~ AstonMartin_col,
                            team == 'Sauber' ~ Kick_col,
                            team == 'Haas F1 Team' ~ Haas_col,
                            team == 'RB F1 Team' ~ RB_col,
                            team == 'Williams' ~ Williams_col,
                            team == 'Alpine F1 Team' ~ Alpine_col)) 
ZeroConstruct <- Constructors2024 %>% filter(round == 1) %>%
  mutate(race = 'Start',
         round = 0,
         points = 0,
         standings = 0,
         date = as.Date('2024-01-01'))
 
Constructors2024 <- rbind(ZeroConstruct, Constructors2024) 

ojs_define(Constructors2024 = Constructors2024 %>% select(round, standings, colour) %>% rename(points = standings))
ojs_define(Constructors = Constructors2024 %>% select(round, standings, team,  colour))
```

`ojs` doesn't use column structured data, but row structured data. To help with this, I've only passed a few columns to the `ojs` variable, and then used `transpose()` to pivot them into the correct format

Now I can build a chart with this data.

I've created a `color` object that lets me control the legend for the graph.

```{ojs}

firstData = transpose(Constructors2024)

 Plot.plot({
    color: {
      legend: true,
      className: "chart-legend",
      domain: transpose(Constructors2024).map(d => d.team), // List of unique categories
      range: transpose(Constructors2024).map(d => d.colour), // Corresponding colors for each category
      legendPosition: "top-right", // Position of the legend
      rows: 2, // Layout: Arrange the legend into 2 rows
      columns: 5,
      swatchWidth: 25,
    },
    style: {
      font: "14px Poppins", // Axis labels, title, etc.
    },
    x: {
      domain: [0, 12], // Set the x-axis range
    },
    y: {
      domain: [0, 350], // Set the y-axis range
    },
    marks: [
      Plot.line(firstData, {
        x: "round", // x-axis row
        y: "points", // y-axis row
        stroke: "team", // Use the category for stroke color
        tip: true, // Enable tooltips
      }),
    ],
  });
```

To style the legend I need to use some CSS:

```{=html}
<style>
.chart-legend-swatch-label {
  font-size: 16px;
  font-family: Poppins;
  font-weight: 400;
}
</style>
```
Now I want to do the same graph, but draw it while we scroll...

```{ojs}
//|eval: false
angleScale2 = d3.scaleLinear()
  .domain([0, 12.1])
  .range([0, 12])
  .clamp(true)

rounds = angleScale2(
  (crTriggerIndex != null ? crTriggerIndex : -1)
    + crTriggerProgress)
```

::: cr-section
::: {#cr-graphic}
```{ojs}
//| eval: false
//| echo: false
d3 = require("d3@7")

reactive_data = transpose(Constructors2024)

rounds = angle2;

viewof reactiveChart = {
  const chart = Plot.plot({
    color: {
      legend: true,
      className: "chart-legend",
      domain: transpose(Constructors2024).map(d => d.team), // List of unique categories
      range: transpose(Constructors2024).map(d => d.colour), // Corresponding colors for each category
      legendPosition: "top-right", // Position of the legend
      rows: 2, // Layout: Arrange the legend into 2 rows
      columns: 5,
      swatchWidth: 25,
    },
    style: {
      font: "14px Poppins", // Axis labels, title, etc.
    },
    x: {
      domain: [0, 12], // Set the x-axis range
    },
    y: {
      domain: [0, 350], // Set the y-axis range
    },
    marks: [
      Plot.line(reactive_data.filter(d => d.round <= rounds), {
        x: "round", // x-axis row
        y: "points", // y-axis row
        stroke: "team", // Use the category for stroke color
        tip: true, // Enable tooltips
      }),
    ],
  });

  // Apply D3 transitions after the chart is rendered
  requestAnimationFrame(() => {
    const svg = d3.select(chart).select("svg"); // Select the SVG of the plot

    // Check if paths exist and print their count
    const paths = svg.selectAll("path").nodes();
    console.log("Number of path elements:", paths.length);

    if (paths.length === 0) {
      console.warn("No path elements found. Verify that the lines were rendered correctly.");
      return; // Exit if no paths are found
    }

    // Apply the animation to all path elements
    svg
      .selectAll("path")
      .attr("stroke-dasharray", function () {
        const totalLength = this.getTotalLength(); // Get the total length of the path
        return `0, ${totalLength}`; // Start with the line hidden
      })
      .transition()
      .duration(2000) // Duration of the transition (2 seconds)
      .attr("stroke-dasharray", function () {
        const totalLength = this.getTotalLength(); // Get the total length of the path
        return `${totalLength}, 0`; // Animate the line being drawn
      });
  });

  return chart; // Return the rendered chart
}


```

```{ojs}
//| echo: false

d3 = require("d3@7")

data = transpose(Constructors)

angleScale2 = d3.scaleLinear()
  .domain([0, 12.1])
  .range([0, 12])
  .clamp(true)

rounds = angleScale2(
  (crTriggerIndex != null ? crTriggerIndex : -1)
    + crTriggerProgress)

function drawChart(data, rounds) {
  // define the graph height and width
  const height = 500;
  const width = 900;
  // define the margins
  const margin = ({top: 20, right: 20, bottom: 30, left: 40});
  
  // Set the position of the x axis 
  const x = d3.scaleLinear() // choose a scale
  .domain(d3.extent(data, d => d.round)) // set the domain, d3.extent pulls the min and max, so here we pull the min and max of round from data
  .range([margin.left, width - margin.right]); // set the range, the physical start and end of the axis. We can use the graph size and margins to do this
  
  const y = d3.scaleLinear()
  .domain(d3.extent(data, d => d.standings))
  .range([height-margin.bottom, margin.top]);
  
  // Declare the line generator.
  const line = d3.line()
    .x(d => x(d.round))
    .y(d => y(d.standings));  
  
  //create the svg that will hold the chart
  const svg = d3.create("svg")
    .attr("width", width)
    .attr("height", height)
    .attr("viewBox", [0,0, width, height])
    .attr("style", "max-width: 100%, height: auto; height: intrinsic;"); //some css for the svg
  
  //add the x-axis to the svg  
  svg.append("g") 
    .attr("transform", `translate(0, ${height - margin.bottom})`)
    .call(d3.axisBottom(x).ticks(width/80).tickSizeOuter(0))
    .call(g => g.select(".domain").remove());
  
// Add the y-axis, remove the domain line, add grid lines and a label
  svg.append("g")
      .attr("transform", `translate(${margin.left},0)`)
      .call(d3.axisLeft(y)
        .ticks((y.domain()[1] - y.domain()[0]) / 50))
      .call(g => g.select(".domain").remove())
      .call(g => g.selectAll(".tick line").clone()
      .attr("x2", width - margin.left - margin.right)
      .attr("stroke-opacity", 0.1))
      .call(g => g.append("text")
      .attr("x", -margin.left)
      .attr("y", 10)
      .attr("fill", "currentColor")
      .attr("text-anchor", "start")
      .text("↑ Points"));  
  
  const groupedData = d3.groups(data, d => d.team)
    .filter(([team, group]) => group.some(d => d.round <= rounds));
    
  // Append a path for the line. Since I have grouped data, this will be done with a for
  groupedData.forEach(([team, group]) => {
   const path = svg.append("path")
    .datum(group)
    .attr("fill", "none")
    .attr("stroke", group[0].colour) 
    .attr("stroke-width", 1.5)
    .attr("d", line);

  // Ensure the path is part of the DOM before animating
  const totalLength = path.node().getTotalLength();

  path
    .attr("stroke-dasharray", totalLength)
    .attr("stroke-dashoffset", totalLength)
    .transition() // Start the transition
    .delay(200)
    .duration(1000) // Duration of 1 second
    .ease(d3.easeCubicInOut) // Smooth easing function
    .attr("stroke-dashoffset", 0); // Reveal the path gradually
  }); 
  
  return svg.node();
  
}

visibility().then(() => {
  // Call drawChart with the latest rounds value whenever visibility is ensured
  drawChart(data, rounds);
});

```

:::

::: progress-block
::: {focus-on="cr-graphic"}
## Round 0

After a dominant 2023 season from Red Bull, everyone hoped for a change in the pecking order.

```{r}
Round <- function(round_no) {
  Round <- Constructors2024 %>% 
    ungroup() %>% 
    filter(round == round_no) %>% 
    select(team, points, standings) %>% 
    mutate(points = paste('+', points), 
           position = row_number())
  gt(Round, rowname_col = 'position')
}

GPName <- function(roundNo) {
  paste0(Constructors2024[Constructors2024$round == roundNo,]$race[roundNo])
}
```
:::

::: {focus-on="cr-graphic"}
## Round 1 - `r GPName(1)`

The `r GPName(1)` didn't look like a promising start...

`r Round(1)`
:::

::: {focus-on="cr-graphic"}
# Round 2 - `r GPName(2)`

`r Round(2)`
:::

::: {focus-on="'cr-graphic'"}
## Round 3 - `r GPName(3)`

After qualifying on pole, Verstappen looked likely to take a record-equalling 10th victory in a row.

However, car problems lead to a surprise retirement on lap 5, opening the door to a 1-2 finish from Ferrari.

`r Round(3)`
:::

Round 4 @cr-graphic

Round 5 @cr-graphic

Round 6 @cr-graphic

Round 7 @cr-graphic

Round 8 @cr-graphic

Round 9 @cr-graphic

Round 10 @cr-graphic

Round 11 @cr-graphic

Round 12 @cr-graphic
:::
:::

```{=html}
<!-- some quick style customisation: see https://closeread.dev/guide/styling -->
<style>
  .cr-section.overlay-left {
    background-color: #ffffff;
  }

  .cr-section.overlay-left .narrative-col .narrative {
    font-size: 1em;
    background-color: #FFFFFF;
    opacity: 1;
    color: #000000;
    padding: 0.6em 0.8em;
    border-radius: 0;
    border: 2px solid #000000;
    box-shadow: 4px 4px 0px 0px #e6e6e6;
    backdrop-filter: blur(10px);
  }
</style>
```
